---
title: "Lab_10"
author: "Yakovenko Ivan"
date: "12/29/2020"
output: 
  rmarkdown::github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load data
Load dat file from hyper

```{r}
df <- read.table("https://hyper.mephi.ru/assets/courseware/v1/63356502d843170fc9f9a568c617f11d/asset-v1:MEPhIx+CS712DS+2020Fall+type@asset+block/jj.dat", header = FALSE, col.names = c('EPS'))
head(df)
```

```{r, echo=FALSE}
if (!require('ggplot2'))
{
  install.packages('ggplot2', dependencies = TRUE)
  library('ggplot2')
}
if (!require('forecast'))
{
  install.packages('forecast', dependencies = TRUE)
  library('forecast')
}
```

## a. After obtaining a copy of the dataset, plot the quarterly EPS vs. time.  Describe any patterns that you observe.
We get nearly exp pattern. time series is growing up.
```{r}
df$ts <- ts(df$EPS, start=1960,frequency=4)
plot(df$ts, xlab = "Time (quaters)",
ylab = "EPS (earnings per share)")
```

## b. In order to perform an ARIMA model, the time series will need to be transformed to remove any trend.  Plot the difference of xt and xt-1, for all t > 0.  

But this trend of growing doesn't describe as stationary series which are needed for ARIMA. For stabilization of data we use "diff" = x(t) - x(t+1) starts from t=1. Difference called Lag.

```{r}
plot(diff(df$ts), xlab = "Time (quaters)",
ylab = "EPS (difference)")
```

We almost detrended our series but its still not so stable
And variance isnt const.

## c. Plot the log10 of the quarterly EPS vs. time and plot the difference of log10(xt ) and log10(xt-1) for all t > 0.  Has this adequately detrended the series?  Has the variability of the differenced log10(EPS) become more constant?  

```{r}
plot(diff(log10(df$ts)), xlab = "Time (quaters)",
ylab = "EPS (log10(difference))")
```

And now our plot is more normalized with log.
Variance is still not same on every year but it seems more stable.

## d. Treating the differenced log10 of the EPS series as a stationary series, plot the ACF and PACF of this series.  What possible ARIMA models would you consider and why? 

Let's plot ACF and PACF to count ARIMA level.

```{r}
par(mfrow=c(2,1))
acf(diff(log10(df$ts))) 
pacf(diff(log10(df$ts)))
```
We see peaks are decaying to 0. So here is AR model. PACF says that order is 1.(p=1) Here we see no zero peaks so no MA model.(q=0)
Or we have

d = 0 because we use diff model.

Here we see peaks on ACF every 4 count
Therefore period=4

● P is the number of terms in the AR model across the s periods.
● D is the number of differences applied across the s periods.
● Q is the number of terms in the MA model across the s periods.

so we get (p,d,q)(P,D,Q)[s]


## e. Run the proposed ARIMA models from part d and compare the results.  Identify an appropriate model.  Justify your choice.

Models are 
(1,0,0)(1,0,0)4
(2,0,1)(1,0,1)4 - means that we use some MA model with this one


```{r}
arima_1 <- Arima(log10(df$ts), order=c(1,0,0),
seasonal = list(order=c(1,0,0),period=4))
arima_1
```

```{r}
arima_2 <- Arima(log10(df$ts), order=c(2,0,1),
seasonal = list(order=c(1,0,1),period=4))
arima_2
```

```{r}
par(mfrow=c(2,1))
Acf(arima_1$residuals) 
pacf(arima_1$residuals)
```
```{r}
par(mfrow=c(2,1))
Acf(arima_2$residuals) 
pacf(arima_2$residuals)
```

So we have AIC lower with (2,0,1)(1,0,1)4 and ACF PACF almost fully in 95%. This model is more appropriate. But it has higher SE so may be it is not the best fit model.


# 4.9

In some cases log isnt the best method to detrend. in general we can use yt = yt - yt+s where s is season.

# 4.10

If we count correlation between indepv we get 1. correlation between num itself also equal to 1.

# 4.11 AR(p), p=1,2,3,4

```{r}
AR4 <- arima.sim(n = 10000, list(ar = c(0.9, -0.5, .2, -.3)))
ts.plot(AR4)
par(mfrow=c(2,1))
acf(AR4)
pacf(AR4)
```

Here we see ACF is more like $sin$. Max on 4

```{r}
AR3 <- arima.sim(n = 10000, list(ar = c(0.9, -0.5, .2)))
ts.plot(AR3)
par(mfrow=c(2,1))
acf(AR3)
pacf(AR3)
```

Max on 3

```{r}
AR2 <- arima.sim(n = 10000, list(ar = c(0.9, -0.5)))
ts.plot(AR2)
par(mfrow=c(2,1))
acf(AR2)
pacf(AR2)
```

Max on 2

```{r}
AR1 <- arima.sim(n = 10000, list(ar = c(0.9)))
ts.plot(AR1)
par(mfrow=c(2,1))
acf(AR1)
pacf(AR1)
```

Max on 1

#4.12 MA(q) q=1,2,3,4

```{r}
MA4 <- arima.sim(n = 10000, list( ma = c(-1.9, 1.7, -1.5, 1.5)))
ts.plot(MA4)
par(mfrow=c(2,1))
acf(MA4)
pacf(MA4)
```

And here is the same - max on 4

```{r}
MA3 <- arima.sim(n = 10000, list( ma = c(-1.9, 1.7, -1.5)))
ts.plot(MA3)
par(mfrow=c(2,1))
acf(MA3)
pacf(MA3)
```
Here we see seasonality. max on 3


```{r}
MA2 <- arima.sim(n = 10000, list( ma = c(-1.9, 1.7)))
ts.plot(MA2)
par(mfrow=c(2,1))
acf(MA2)
pacf(MA2)
```
 max on 2

```{r}
MA1 <- arima.sim(n = 10000, list( ma = c(-1.9)))
ts.plot(MA1)
par(mfrow=c(2,1))
acf(MA1)
pacf(MA1)
```

And max on 1

