---
title: "lab_23"
author: "Yakovenko Ivan"
date: "11/9/2020"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Analysis in R

## Load Data

First we need to load our data from file. Let's see what we get:

```{r}
zipIncome <- read.table('https://hyper.mephi.ru/assets/courseware/v1/94f633ca057a1aa84db0364cf4bfa81d/asset-v1:MEPhIx+CS712DS+2020Fall+type@asset+block/zipIncome.txt', header = TRUE, sep = '|', col.names = c('zipCode', 'income'), comment.char = "(")

head(zipIncome, n = 10)
```

## Analysis

To find median and mean of average **income** we use `summary()`

```{r, results='asis'}
tab_summary <- summary(zipIncome)
cat("**For income:**", tab_summary[,2], sep="\n")
```

```{r}
mean(zipIncome$income)
x <- zipIncome$income
density(x)$x[which.max(density(x)$y)]
plot(density(x)$y)

```


Let's look at a scatter plot of data. Outlier values are $\$0$ and $\$250000$

```{r}
plot(zipIncome)
```

To avoid outliers let's create subset of data where $\$7,000 < income < \$200,000$ and build a simple boxplot:

```{r}
subset_data <- subset(zipIncome, income>7000 & income < 200000)
tab_subset <- summary(subset_data)
cat("**For income:**", tab_subset[,2], sep="\n")
```

```{r}
boxplot(formula=income ~ zipCode, col="white", data = subset_data, main = 'Average Household Income by Zip Code', xlab = 'Zip Code', ylab = 'Income')
```


**Or with log scale over Y**

```{r}
boxplot(income ~ zipCode, col="white", data = subset_data, main = 'Average Household Income by Zip Code', xlab = 'Zip Code', ylab = 'Income', log = 'y')
```

```{r}
library(lattice)
densityplot(log(zipIncome$income[zipIncome$zip==0]))
densityplot(zipIncome$income[zipIncome$zip==0])
```


For the next step we are needed in **ggplot2** package. Let's check it and install if it is necessary.

```{r, results='hide'}
 if (!require('ggplot2'))
   {
      install.packages('ggplot2', dependencies = TRUE)
      library('ggplot2')
   }
```


### Making plots with ggplot2

```{r}
plt <- ggplot(data=subset_data, aes(as.factor(zipCode), income))
plt + geom_point(position='jitter', alpha=0.2) + scale_y_log10()
```


And colorful ggplot:

```{r}
color_plt <- ggplot(data=subset_data, aes(as.factor(zipCode), income))
color_plt <- color_plt + geom_point(aes(color=factor(zipCode)), position='jitter', alpha=0.2) + scale_y_log10()
color_plt <- color_plt + geom_boxplot(aes(group=zipCode), alpha = 0.1, outlier.size = -Inf)
plot(color_plt + labs(title = 'Average Household Income by Zip Code', x = 'Zip Code', y = 'Income', color = 'Region'))
```

## Conclusion

After basic analysis of this data we can say that 0 and 9 zip codes have the largest average income.
